### 2025-04-26 EDA/multicollinearity_and_outlier.qmd
Initial Commit

---
title: "outlier_exploration"
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(ggthemes)
library(car)
```

```{r}
full_model <- lm(sleep_duration ~., sleep)
```

### Outliers

```{r}
plot(full_model)
```

```{r}
leverage <- hatvalues(full_model)
studentized <- rstudent(full_model)
dffits_vals <- dffits(full_model)
cooks_d <- cooks.distance(full_model)

diagnostics <- data.frame(Leverage = leverage,
                          Studentized = studentized,
                          DFFITS = dffits_vals,
                          Cooks_Distance = cooks_d)

n <- nrow(sleep)
p <- length(coef(full_model)) - 1
leverage_threshold <- 2 * (p + 1) / n
studentized_threshold <- 2
cooks_threshold <- 1

# Find points exceeding thresholds
high_leverage <- which(leverage > leverage_threshold)
high_studentized <- which(abs(studentized) > studentized_threshold)
high_cooks <- which(cooks_d > cooks_threshold)

# Unique indices that exceed any threshold
suspicious_indices <- sort(unique(c(high_leverage, high_studentized, high_cooks)))

# Return threshold values, diagnostic preview, and rows flagged as potential outliers
list(
  leverage_threshold = leverage_threshold,
  studentized_threshold = studentized_threshold,
  cooks_threshold = cooks_threshold,
  diagnostics_preview = head(diagnostics, 10),
  flagged_rows = diagnostics[suspicious_indices, ])

```

### Multicollinearity