### 2025-04-26 EDA/multicollinearity_and_outlier.qmd
Ethan: Removed aliased variable (BMI), obtained VIF values to observe multicolinearity, used step-wise AIC variable selection which removed quality_of_sleep variable, obtained VIF values of new model and were improved dramatically.

---
title: "outlier_exploration"
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(ggthemes)
library(car)
```

```{r}
full_model <- lm(sleep_duration ~., sleep)
summary(full_model)
```

### Outliers

```{r}
plot(full_model)
```

```{r}
leverage <- hatvalues(full_model)
studentized <- rstudent(full_model)
dffits_vals <- dffits(full_model)
cooks_d <- cooks.distance(full_model)

diagnostics <- data.frame(Leverage = leverage,
                          Studentized = studentized,
                          DFFITS = dffits_vals,
                          Cooks_Distance = cooks_d)

n <- nrow(sleep)
p <- length(coef(full_model)) - 1
leverage_threshold <- 2 * (p + 1) / n
studentized_threshold <- 2
cooks_threshold <- 1

# Find points exceeding thresholds
high_leverage <- which(leverage > leverage_threshold)
high_studentized <- which(abs(studentized) > studentized_threshold)
high_cooks <- which(cooks_d > cooks_threshold)

# Unique indices that exceed any threshold
suspicious_indices <- sort(unique(c(high_leverage, high_studentized, high_cooks)))

# Return threshold values, diagnostic preview, and rows flagged as potential outliers
list(
  leverage_threshold = leverage_threshold,
  studentized_threshold = studentized_threshold,
  cooks_threshold = cooks_threshold,
  diagnostics_preview = head(diagnostics, 10),
  flagged_rows = diagnostics[suspicious_indices, ])

# Remove the suspicious rows
sleep_clean <- sleep[-suspicious_indices, ]

# Refit the model excluding suspicious rows 
full_model_clean <- lm(sleep_duration ~ ., data = sleep_clean)

# Plot diagnostics again on the cleaned model
par(mfrow = c(2, 2))
plot(full_model_clean)

# Comparison of model summaries
summary(full_model)        # Old model (with potential outliers)  R^2  = 0.9592
summary(full_model_clean)  # New model (without potential outliers)  R^2  = 0.9929

```

```{r}

```

### Multicollinearity

```{r}
library(car)

# Calculate VIFs on the cleaned model
vif_values <- vif(full_model_clean)

# Running VIF gave error (aliased variables are present in the model), so find out which are aliased 
alias(full_model_clean) # BMI->Overweight is perfectly aliased with other variables in the model

# Refit the model without BMI
full_model_clean2 <- lm(sleep_duration ~ gender + age + occupation + quality_of_sleep +
                        physical_activity_level + stress_level + heart_rate +
                        daily_steps + sleep_disorder + MAP, data = sleep_clean)

# Calculate VIFs on the cleaned model, with BMI removed 
vif_values <- vif(full_model_clean2)
vif_values     # Significant multicolinearity present, variable selection is next. Start with step-wise AIC variable selection

```

### Variable Selection

```{r}
# Step-wise AIC Variable Selection

library(MASS)

step_model <- stepAIC(full_model_clean2, direction = "both", trace = FALSE)

# Check final selected model
summary(step_model)

# Observing which variables were removed from full model 
formula(full_model_clean2) 
formula(step_model)      # Variable removed: quality_of_sleep

# Check VIF of the step-wise model
vif(step_model)      # VIF values have decreased significantly after step-wise variable selection

```