### 2025-05-07 Kyubin's Edit/KYUBIN_EDIT.qmd
Lines 96-194 were edited: Added in the Code taken from "multicolinearity_and_outlier.qmd" that justifies obs 264 is an outlier Created the new dataset that removed obs 264 called sleep_test Created the new full model that removed obs 264 called full_model_test Changed "sleep" to "sleep_test" for for the VIF testing

---
title: "outlier_exploration"
format: pdf
---

```{r}
library(readr)
library(tidyverse)
library(janitor)
library(dplyr)

sleep <- read_csv("../data/Sleep_health_and_lifestyle_dataset.csv") %>%  
  clean_names()

sleep <- sleep %>%
  separate(blood_pressure, into = c("systolic", "diastolic"), sep = "/", convert = TRUE)

sleep$stress_level <- as.factor(sleep$stress_level)
sleep$gender <- as.factor(sleep$gender)
sleep$occupation <- as.factor(sleep$occupation)
sleep$bmi_category <- as.factor(sleep$bmi_category)
sleep$sleep_disorder <- as.factor(sleep$sleep_disorder)

sleep <- sleep %>%
  mutate(systolic = as.numeric(systolic),
         diastolic = as.numeric(diastolic),
         PP = systolic-diastolic)

sleep <- sleep %>%
  dplyr::select(!c(person_id, systolic, diastolic))

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(ggthemes)
library(car)
```

```{r}
full_model <- lm(sleep_duration ~., sleep)
summary(full_model)
```

### Outliers

```{r}
plot(full_model)
```

```{r}
# Calculate Cook's Distance
cooks_d <- cooks.distance(full_model)
summary(cooks_d)

# Threshold
n <- nrow(sleep)
cooks_threshold <- 4 / n

# Basic Cook's Distance plot
plot(cooks_d, type = "h", 
     main = "Cook's Distance", 
     ylab = "Cook's D", 
     xlab = "Observation", 
     col = "blue")

# Add horizontal threshold line
abline(h = cooks_threshold, col = "red", lty = 2)

# Label points above threshold
high_cooks <- which(cooks_d > cooks_threshold)

# Add text labels
text(x = high_cooks, 
     y = cooks_d[high_cooks], 
     labels = high_cooks, 
     pos = 3, 
     cex = 0.8, 
     col = "red")

# Found that obs 6, 7, 107 were pretty large

```

```{r}
# Look into obs 6, 7, 107
sleep[6,]
sleep[7,]
sleep[107,]

# They all seem pretty valid
```

### Code taken from "multicolinearity_and_outlier.qmd" that justifies obs 264 is an outlier

```{r}
leverage <- hatvalues(full_model)
studentized <- rstudent(full_model)
dffits_vals <- dffits(full_model)
cooks_d <- cooks.distance(full_model)

diagnostics <- data.frame(Leverage = leverage,
                          Studentized = studentized,
                          DFFITS = dffits_vals,
                          Cooks_Distance = cooks_d)

n <- nrow(sleep)
p <- length(coef(full_model)) - 1
leverage_threshold <- 2 * (p + 1) / n
studentized_threshold <- 2
cooks_threshold <- 1

# Find points exceeding thresholds
high_leverage <- which(leverage > leverage_threshold)
high_studentized <- which(abs(studentized) > studentized_threshold)
high_cooks <- which(cooks_d > cooks_threshold)

# Unique indices that exceed any threshold
suspicious_indices <- sort(unique(c(high_leverage, high_studentized, high_cooks)))

# Return threshold values, diagnostic preview, and rows flagged as potential outliers
list(
  leverage_threshold = leverage_threshold,
  studentized_threshold = studentized_threshold,
  cooks_threshold = cooks_threshold,
  diagnostics_preview = head(diagnostics, 10),
  flagged_rows = diagnostics[suspicious_indices, ])

# observation 264 has a leverage of 1

# New data set used for model
sleep_test <- sleep[-264,]

full_model_test<-lm(sleep_duration ~., sleep_test)
summary(full_model_test)
```

### Multicollinearity

```{r}
library(car)

vif(full_model_test)

#GVIF high for occupation, stress level, bmi category
#GVIF^ high for quality of sleep

#stress
sleep_test$stress_level_num <- as.numeric(as.factor(sleep_test$stress_level))

aux_model_stress_level <- lm(stress_level_num ~ gender + age + quality_of_sleep + 
                    physical_activity_level + occupation + 
                    bmi_category + heart_rate + daily_steps + 
                    sleep_disorder + PP, data = sleep_test)
summary(aux_model_stress_level)
# R^2 .9563 suggest we remove it

#occupation
sleep_test$occupation_num <- as.numeric(as.factor(sleep_test$occupation))

aux_model_occupation <- lm(occupation_num ~ gender + age + quality_of_sleep + 
                           physical_activity_level + bmi_category + heart_rate + 
                           daily_steps + sleep_disorder + PP + stress_level, 
                           data = sleep_test)

summary(aux_model_occupation)
# R^2 around .7453, keep occupation

# BMI
sleep_test$bmi_category_num <- as.numeric(as.factor(sleep_test$bmi_category))

aux_model_bmi <- lm(bmi_category_num ~ gender + age + quality_of_sleep + 
                     physical_activity_level + occupation + heart_rate + 
                     daily_steps + sleep_disorder + PP + stress_level,
                     data = sleep_test)

summary(aux_model_bmi)
# R^2 around 0.9326

#quality
sleep_test$quality_of_sleep_num <- as.numeric(as.factor(sleep_test$quality_of_sleep))

aux_model_quality <- lm(quality_of_sleep_num ~ gender + age + physical_activity_level +
                         occupation + bmi_category + heart_rate + daily_steps +
                         sleep_disorder + PP + stress_level,
                         data = sleep_test)

summary(aux_model_quality)
# R^2 around 0.9673

```

## Table of R^2 Auxiliary Model
```{r}
library(dplyr)

# Define the full list of predictors (excluding the outcome)
predictors <- c("stress_level", "gender", "age", "quality_of_sleep",
                "physical_activity_level", "occupation", "bmi_category",
                "heart_rate", "daily_steps", "sleep_disorder", "MAP")

# Create a function to fit auxiliary model and return R² and Adjusted R²
get_aux_model_r2 <- function(var, data) {
  # Convert response to numeric (safely)
  data[[paste0(var, "_num")]] <- as.numeric(as.factor(data[[var]]))

  # Define predictors excluding the variable of interest
  others <- setdiff(predictors, var)
  formula <- as.formula(paste(paste0(var, "_num"), "~", paste(others, collapse = " + ")))

  # Fit auxiliary model
  model <- lm(formula, data = data)

  # Extract R² and Adjusted R²
  r2 <- summary(model)$r.squared
  adj_r2 <- summary(model)$adj.r.squared

  return(data.frame(Variable = var, R_squared = r2, Adjusted_R_squared = adj_r2))
}

# Apply the function to all predictors
aux_results <- do.call(rbind, lapply(predictors, get_aux_model_r2, data = sleep))

# Print results
print(aux_results)
```

```{r}

# get rid of 3 highest aux r^2 adj and high vif scores: BMI and Quality
model_vif <- lm(sleep_duration ~gender + age + stress_level +  
                    physical_activity_level + occupation + 
                    heart_rate + daily_steps + 
                    sleep_disorder + MAP, data = sleep)

# View the summary
summary(model_vif)

vif(model_vif)
```

### Variable Selection w VIF
```{r}

# 2. Backward Elimination
model_backward <- step(model_vif, direction = "backward")
summary(model_backward)

# 3. Forward Selection
# Define the full scope and a null model
full_scope <- formula(model_vif)
null_model <- lm(sleep_duration ~ 1, data = sleep)

model_forward <- step(null_model, scope = full_scope, direction = "forward")
summary(model_forward)

# 4. Stepwise Selection (both directions)
model_stepwise <- step(null_model, scope = full_scope, direction = "both")
summary(model_stepwise)

```

# Lasso after VIF and Variable Selection
```{r}

library(glmnet)
library(tidyverse)
library(tidymodels)
library(tidyr)
library(car)
library(caret)

sleep <- subset(sleep, select = -c(bmi_category, quality_of_sleep))

## Split Training/Test Data
sleep_sp <- sleep %>% initial_split(prop = 0.8, strata=sleep_duration)
rtrain <- sleep_sp %>% training()
rtest <- sleep_sp %>% testing()

model0 <- lm(sleep_duration ~ ., data=rtrain)
summary(model0)

## Regularization
xrtrain <- rtrain %>% dplyr::select(-sleep_duration) %>% as.matrix()
yrtrain <- rtrain$sleep_duration

xrtest  <- rtest %>% dplyr::select(-sleep_duration) %>% as.matrix()
yrtest  <- rtest$sleep_duration

modelL <- glmnet(xrtrain, yrtrain)
plot(modelL,label = TRUE)
## Df=# of nonzero coefficients, %Dev = % of Deviance explained (like SSR)
print(modelL)
## Coefficient for a certain lambda
coef.glmnet(modelL, s= 0.1)

## Cross-Validation
## nfolds = # of groups ex) 20
## Regression with 19 groups and test with 1 group
cvmodelL <- cv.glmnet(xrtrain, yrtrain, type.measure = "mse", nfolds = 20)
coef.glmnet(modelL, s= cvmodelL$lambda.min)
plot(cvmodelL)

coef_min_lambda <- coef.glmnet(modelL, s = cvmodelL$lambda.min)

# Convert to a tidy data frame and filter out zero coefficients
nonzero_coefs <- as.matrix(coef_min_lambda)
nonzero_coefs_df <- data.frame(
  Variable = rownames(nonzero_coefs),
  Coefficient = nonzero_coefs[, 1]
)

# View coefficients that were shrunk to zero
zeroed_out <- nonzero_coefs_df %>% filter(Coefficient == 0)
print(zeroed_out)

# gender, occupation and sleep disorder zeroed out and age is insignificant

model_LASSO <- lm(sleep_duration ~ stress_level +  
                    physical_activity_level + 
                    heart_rate + daily_steps + 
                    MAP, data = sleep)
summary(model_LASSO)

```
## Ridge if neccessary

```{r}
modelR <- glmnet(xrtrain, yrtrain, alpha=0)
plot(modelR,label = TRUE)
## Df=# of nonzero coefficients, %Dev = % of Deviance explained (like SSR)
print(modelR)
## Coefficient for a certain lambda
coef.glmnet(modelR, s= 0.1) ## s= lambda value (any)

## CrossValidation
cvmodelR <- cv.glmnet(xrtrain, yrtrain, alpha=0, type.measure = "mse", nfolds = 20)
plot(cvmodelR)
log(cvmodelR$lambda.min)
```