### 2025-05-08 paper/paper.qmd
adding in table for coefficients of final model

---
title: "Multiple Linear Regression on Sleep Data"
author:
  - name: Ethan Newcomb
  - name: Kyubin Im
  - name: Matthew Murnane
format:
  pdf:
    mainfont: "Times New Roman"
indent: true
fontsize: "12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
#Packages
library(tidyverse)
library(dplyr)
library(ggthemes)
library(tidyr)
library(knitr)
library(kableExtra)
library(car)
library(MASS)
library(readr)
library(janitor)
library(patchwork)
library(broom)
library(glmnet)
library(caret)
library(tidymodels)
library(rsample)
```

```{r include=FALSE}
#Data Wrangling Code
sleep <- read_csv("../data/Sleep_health_and_lifestyle_dataset.csv") %>%  
  clean_names()

sleep <- sleep %>%
  separate(blood_pressure, into = c("systolic", "diastolic"), sep = "/", convert = TRUE)

sleep$stress_level <- as.factor(sleep$stress_level)
sleep$gender <- as.factor(sleep$gender)
sleep$occupation <- as.factor(sleep$occupation)
sleep$bmi_category <- as.factor(sleep$bmi_category)
sleep$sleep_disorder <- as.factor(sleep$sleep_disorder)

sleep <- sleep %>%
  mutate(systolic = as.numeric(systolic),
         diastolic = as.numeric(diastolic),
         PP = systolic-diastolic)

sleep <- sleep %>%
  dplyr::select(!c(person_id, systolic, diastolic))
```

\newpage

\begin{center}
\Large\textbf{\large Introduction}
\end{center}

\noindent\textbf{Motivation}

\vspace{0.5em}

Sleep is an important contributor to our well-being and is deeply connected to our physical health, and cognitive function. However we often find that many individuals struggle with getting long hours sleep, often finding themselves tossing and turning throughout the night. Consequently, there is now a growing interest in understanding the lifestyle and bio metric factors that influence sleep patterns and duration. General knowledge and intuition suggest factors such as physical activity and occupational stress may influence sleep duration. Therefore, our goal is to use linear regression techniques to quantify exactly how much these factors reduce or extend sleep time.

Before, analyzing the data set, we expect a number of our indicators to contribute significantly to sleep duration. In general, we would assume that occupation will play an important role in analyzing our data as a more stressful work environment may hinder sleep patterns and disturb the body's natural sleep cycle. Additionally, we may expect that higher levels of physical activity will work in tandem with deeper and longer duration of sleep.

Also, we seek to filter through and find the most important variables in our analysis. We expect that variables like Daily Steps and Physical Activity Level will be a contribute to sleep duration in a similar manner. In our data set, we have numerous combinations of variables that are alike in this way and therefore we want to reduce our factors to give the most precise and simple model possible.

\vspace{1.5em}

\noindent\textbf{About the Data}

```{r}
#Table 1: Variables in Sleep
variable_names <- matrix(c(
  "Gender", "Age", "Occupation",
  "Sleep Duration (hours)", "Quality of Sleep (scale: 1–10)", "Physical Activity Level (minutes/day)",
  "Stress Level (scale: 1–10)", "BMI Category", "PP (Pulse Pressure)",
  "Heart Rate (bpm)", "Daily Steps", "Sleep Disorder"
), ncol = 3, byrow = TRUE)

kable(variable_names, col.names = NULL, caption = "Variables From Our Data")
```

Sleep Duration is a continuous variable and will be our response variable. The rest will be our covariates. Gender, Occupation, Quality of Sleep, Stress Level, BMI Category and Sleep Disorder are categorical variables. Physical Activity, PP, agem and Daily steps are continuous variables. We have a total of 374 observations.

We had no missing data in the data set but we did have to make a mutation using `dplyr`. In the original data set there was a blood pressure variable that was encoded as a character but was a ratio, e.g. "$\frac{120}{80}$". We separated the blood pressure variable into two: Systolic and Diastolic. These two we suspected to be highly correlated but did not want to just drop one or the other. We had options of combining the two into something meaningful. One was Pulse Pressure (PP), which is calculated as $\text{PP}=\text{systolic}-\text{diastolic}$. PP is the force of the hearts contraction on the arteries. The other is called Mean Arterial Pressure (MAP), which is calculated as $\text{MAP}=\frac{\text{systolic}\cdot\text{diastolic}^2}{3}$. MAP is the average blood pressure throughout a cardiac cycle. We decided on PP because it was simpler to calculate and understand.

For transparency it should be noted that the data is synthetic. Even though it is synthetic we still treat the analysis and report as if it was real data.

\noindent\textbf{Exploratory Data Analysis}

```{r}
p1 <- sleep %>% 
  ggplot() +
  geom_jitter(aes(x = age, y = sleep_duration), size = .5) +
  theme_few() +
  labs(x = "Age", y = NULL)

p2 <- sleep %>% 
  ggplot() +
  geom_jitter(aes(x = PP, y = sleep_duration), size = .5) +
  theme_few() +
  labs(x = "Pulse Pressure", y = NULL)

p3 <- sleep %>% 
  ggplot() +
  geom_jitter(aes(x = heart_rate, y = sleep_duration), size = .5) +
  theme_few() +
  labs(x = "Heart Rate", y = NULL)

p4 <- sleep %>% 
  ggplot() +
  geom_jitter(aes(x = daily_steps, y = sleep_duration), size = .5) +
  theme_few() +
  labs(x = "Daily Steps", y = NULL)

(p1 + p2) / (p3 + p4) +
  plot_annotation(
    title = "Sleep Duration vs Continuous Variables",
    theme = theme(plot.title = element_text(size = 16, hjust = 0.5))
  ) &
  theme(axis.title.y = element_text(angle = 90)) &
  labs(y = "Sleep Duration")
```
Visual relationship can be seen for `Sleep Duration` with `Age` and `sleep_duration`. Not so much for `PP` and `daily_steps`. The correlation between our numeric covariates is most notable between `sleep_duration` and `heart_rate`. `age` and `hearts_pressure`

```{r}
corr_vars <- sleep %>%
  dplyr::select(sleep_duration, age, PP, heart_rate, daily_steps)

corr_matrix <- round(cor(corr_vars, use = "complete.obs"), 2)

knitr::kable(corr_matrix, caption = "Correlation Matrix of Continuous Varaibles")
```

\vspace{1.5em}

\begin{center}
\Large\textbf{\large Finding Our Model}
\end{center}

\noindent\textbf{Full Model}

Running the full model we can see that our omnibus hypothesis test has an F-stat of 426.6 which is greater than the 99th percentile for our F-distribution, $F_{31,342,.99}=$ `r round(qf(0.99, 31, 342),2)`. So we reject the null that all coefficients are 0. We also see a high Adjusted $R^2$ and very low Residual Sum of Squares compared to the Total Sum of Squares.

```{r}
full_model <- lm(sleep_duration ~ ., data = sleep)
full_model_summary <- summary(full_model)

model_stats <- tibble::tibble(`F-statistic` = round(full_model_summary$fstatistic[1], 2),
                              `Adjusted R²` = round(full_model_summary$adj.r.squared, 3),
                              `Residual Sum of Squares` = round(sum(residuals(full_model)^2), 2),
                              `Total Sum of Squares` = round(sum((sleep$sleep_duration - mean(sleep$sleep_duration))^2), 2))

knitr::kable(model_stats, caption = "Full Linear Model Outcome")
```
We also saw all coefficients have standard error less than one.

```{r}
model_data <- augment(full_model)

p5 <- ggplot(model_data, aes(.fitted, .resid)) +
  geom_point(size = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_few() +
  labs(title = "Residuals vs Fitted",
       x = "Fitted values", y = "Residuals")

p6 <- ggplot(model_data, aes(sample = .resid)) +
  stat_qq(size = 1) +
  stat_qq_line() +
  theme_few() +
  labs(title = "Normal Q-Q Plot", x = "Theoretical Quantiles", y = "Sample Quantiles")

p5 / p6
```

As far as assumptions go homoscedasticity seems to be true and normality might be slightly violated with heavy tails but there is no excessive deviation from normality. We will explore multicollinearity and outliers next.

```{r}
kable(vif(full_model))
```
`age` and `quality_of_sleep` have $\text{GVIF}^{2\cdot\text{DF}} > 5$. We may want to explore Lasso to deal with this.

Looking now towards outliers we selected observations that had one of following: leverage $> 2\cdot\frac{p+1}{n}$, studentized residual > 2, and a cooks threshold > 1. We saw a total of 57 observations that exceeded at least one of these thresholds. Since we can't say that any of these observation are errors we will only remove row 264 because it has a leverage of 1 and `NaN` for the rest. This is clearly a problem observation.

```{r}
#Values of Interest
leverage <- hatvalues(full_model)
studentized <- rstudent(full_model)
dffits_vals <- dffits(full_model)
cooks_d <- cooks.distance(full_model)

diagnostics <- data.frame(Leverage = leverage,
                          Studentized = studentized,
                          DFFITS = dffits_vals,
                          Cooks_Distance = cooks_d)

#Thresh hold
n <- nrow(sleep)
p <- length(coef(full_model)) - 1
leverage_threshold <- 2 * (p + 1) / n
studentized_threshold <- 2
cooks_threshold <- 1

# Find points exceeding thresholds
high_leverage <- which(leverage > leverage_threshold)
high_studentized <- which(abs(studentized) > studentized_threshold)
high_cooks <- which(cooks_d > cooks_threshold)

# Unique indices that exceed any threshold
suspicious_indices <- sort(unique(c(high_leverage, high_studentized, high_cooks)))

aggregation <- list(leverage_threshold = leverage_threshold,
                    studentized_threshold = studentized_threshold,
                    cooks_threshold = cooks_threshold,
                    flagged_rows = diagnostics[suspicious_indices, ])
```

```{r}
aggregation$flagged_rows %>% 
  arrange(desc(Leverage)) %>% 
  head(n = 1) %>% 
  kable(digits = 3, caption = "Row with Highest Leverage")
```

\noindent\textbf{Updating Model: VIF and Lasso Considerations}

We removed variables `age` and `quality_of_sleep` and ran a MLR model. Our results showed that the reduced VIF model was very similar to the full model. Adjusted R^2 was essentially unchanged. Therefore, we concluded that `age` and `quality_of_sleep` did not contribute much to the regression analysis because they were linear combinations of other independent variables.

```{r}
sleep_test <- sleep[-264,]

sleep_test <- sleep_test %>%
  dplyr::select(!c(age, quality_of_sleep))

test_model<-lm(sleep_duration ~., sleep_test)
test_model_summary <- summary(test_model)

test_model_stats <- tibble::tibble(`F-statistic` = round(test_model_summary$fstatistic[1], 2),
                              `Adjusted R²` = round(test_model_summary$adj.r.squared, 3),
                              `Residual Sum of Squares` = round(sum(residuals(test_model)^2), 2),
                              `Total Sum of Squares` = round(sum((sleep_test$sleep_duration-mean(sleep_test$sleep_duration))^2), 2))

knitr::kable(test_model_stats, caption = "MLR Outcome for age and quality of sleep removed")
```

In order to avoid overfitting issues, we utilized a greedy algorithmn process to select variables that provided the lowest AIC value. We ran 3 model selection process as discussed in class and recorded the variables selected for each method. We found that all 3 methods took all of our variables in our data set. We knew that AIC penalizes the fitted model for each additional variables taken. Despite this, we saw that all variables were significant. Note that the variable selection process was done after our high VIF variables were removed.

```{r include=FALSE}
# 1. Model Selection using full_model_vif
model_backward <- stats::step(test_model, direction = "backward")

# Define full model formula and null model for forward/stepwise
full_scope <- formula(test_model)
null_model <- lm(sleep_duration ~ 1, data = sleep_test)

model_forward <- stats::step(null_model, scope = full_scope, direction = "forward")
model_stepwise <- stats::step(null_model, scope = full_scope, direction = "both")

# 2. Extract variable names selected by each model
vars_backward <- attr(terms(model_backward), "term.labels")
vars_forward <- attr(terms(model_forward), "term.labels")
vars_stepwise <- attr(terms(model_stepwise), "term.labels")

# 3. Combine all variables that appeared in any model
all_vars <- sort(unique(c(vars_backward, vars_forward, vars_stepwise)))

# 4. Create a comparison table
selection_table <- data.frame(
  Variable = all_vars,
  Backward = ifelse(all_vars %in% vars_backward, "Yes", "No"),
  Forward  = ifelse(all_vars %in% vars_forward, "Yes", "No"),
  Stepwise = ifelse(all_vars %in% vars_stepwise, "Yes", "No")
)
```

```{r}
kable(selection_table, caption = "Variable Selection")
```

For our suggested regression, we used LASSO regression. We hoped to extract the most import input features to achieve a simple yet powerful model. To find the optimal $\lambda$ or the regularization parameter, we used a cross-validation process by splitting our data set: train = 80%. We saw that `gender`, `occupation`, `bmi_category`, and `sleep_disorder` converged to zero. Optimal lambda value was 0.0009622.

```{r include=FALSE}
set.seed(510)

## Split Training/Test Data
sleep_sp <- sleep_test %>% initial_split(prop = 0.8, strata=sleep_duration)
rtrain <- sleep_sp %>% training()
rtest <- sleep_sp %>% testing()

model0 <- lm(sleep_duration ~ ., data=rtrain)
summary(model0)

## Regularization
xrtrain <- rtrain %>% dplyr::select(-sleep_duration) %>% as.matrix()
yrtrain <- rtrain$sleep_duration

xrtest  <- rtest %>% dplyr::select(-sleep_duration) %>% as.matrix()
yrtest  <- rtest$sleep_duration

modelL <- glmnet(xrtrain, yrtrain)
plot(modelL,label = TRUE)
## Df=# of nonzero coefficients, %Dev = % of Deviance explained (like SSR)
print(modelL)
## Coefficient for a certain lambda
coef.glmnet(modelL, s= 0.1)

## Cross-Validation
## nfolds = # of groups ex) 20
## Regression with 19 groups and test with 1 group
cvmodelL <- cv.glmnet(xrtrain, yrtrain, type.measure = "mse", nfolds = 20)
coef.glmnet(modelL, s= cvmodelL$lambda.min)
plot(cvmodelL)

coef_min_lambda <- coef.glmnet(modelL, s = cvmodelL$lambda.min)

# Convert to a tidy data frame and filter out zero coefficients
nonzero_coefs <- as.matrix(coef_min_lambda)
nonzero_coefs_df <- data.frame(
  Variable = rownames(nonzero_coefs),
  Coefficient = nonzero_coefs[, 1]
)

```

```{r}
kable(nonzero_coefs_df, digits = 4, caption = "Non-Zero Coefficients from LASSO (Selected by Cross-Validation)",
      row.names = FALSE)

# Create a tibble with the optimal lambda
lambda_table <- tibble(
  Description = "Optimal lambda (lambda.min)",
  Value = cvmodelL$lambda.min
)

# Display in kable format
kable(lambda_table, caption = "Optimal Lambda from Cross-Validation")

```

```{r}
sleep_test_final <- sleep_test %>%
  dplyr::select(!c(gender, occupation, bmi_category, sleep_disorder, daily_steps))

final_model <- lm(sleep_duration ~., sleep_test_final)
final_model_summary <- summary(final_model)

final_model_stats <- tibble::tibble(`F-statistic` = round(final_model_summary$fstatistic[1], 2),
                              `Adjusted R²` = round(final_model_summary$adj.r.squared, 3),
                              `Residual Sum of Squares` = round(sum(residuals(final_model)^2), 2),
                              `Total Sum of Squares` = round(sum((sleep_test_final$sleep_duration-mean(sleep_test_final$sleep_duration))^2), 2))

knitr::kable(final_model_stats, caption = "Final Model Statistics")

final_model <- lm(sleep_duration ~ ., data = sleep_test_final)

tidy(final_model) %>%
  dplyr::select(term, estimate, p.value) %>%
  kable(
    digits = 3,
    caption = "Coefficients and P-values from Linear Model",
    booktabs = TRUE,
    format = "latex",
    escape = TRUE
  )

```

\begin{center}
\Large\textbf{\large Discussion}
\end{center}

Upon comparing the final model to the original full model, we observe a reduction in the Adjusted R² from 0.958 to 0.880, and an increase in the Residual Sum of Squares (RSS) from 9.28 to 27.61. The F-statistic increased slightly from 313.34 to 343.56, while the Total Sum of Squares remained nearly unchanged at approximately 236. This suggests that, although the simplified model explains slightly less of the variation in sleep duration, it still performs quite well in terms of overall fit and statistical significance.

Despite the modest drop in Adjusted R², we are confident in selecting the final model. This model is much simpler, yet retains highly significant explanatory power, making it more interpretable and potentially more useful in real-world applications. The use of Lasso regularization, which penalizes more complex models, likely contributed to the decrease in Adjusted R². This trade-off was intentional. We prioritized a simpler model with a more strict variable selection.
This has important practical benefits: a simpler model enables more efficient and cost-effective data collection for future studies. For example, if a healthcare provider wanted to screen individuals for sleep issues, collecting a smaller set of variables would be faster and less resource-demanding.
In the final model, several predictors emerged as statistically significant at the 5% level, including: Physical Activity Level, Stress Levels 4 through 8, and Pulse Pressure (PP).

These results align well with our expectations—individuals experiencing higher stress levels or reduced physical activity tend to sleep less. Interestingly, heart rate, while retained in the model, was not statistically significant at the 5% level, suggesting its role may be less direct in predicting sleepduration.
Lastly, 10-fold cross-validation showed that both the full and final models performed well, with the final model achieving only a slight decrease in R² but with a meaningful gain in simplicity. This further supports our choice to recommend the final model as a strong, interpretable, and efficient tool for predicting sleep duration.

```{r include=FALSE}
library(caret)
library(knitr)
library(dplyr)

set.seed(510)

# Set up 10-fold cross-validation
train_control <- trainControl(method = "cv", number = 10)

# 1. Initial model using full dataset
cv_full <- train(sleep_duration ~ ., data = sleep, method = "lm", trControl = train_control)

# 2. Final model using sleep_test_final dataset
cv_final <- train(sleep_duration ~ ., data = sleep_test_final, method = "lm", trControl = train_control)

# 3. Extract RMSE and R-squared
cv_results <- tibble(
  Model = c("Initial Model", "Final Model"),
  RMSE = c(cv_full$results$RMSE, cv_final$results$RMSE),
  Rsquared = c(cv_full$results$Rsquared, cv_final$results$Rsquared)
)

```

```{r}
# 4. Display as kable table
kable(cv_results, caption = "10-Fold Cross-Validation Results for Initial and Final Models")
```